Distributional Trees and Forests

### 2017-11-04 pkg/disttree/inst/draft/draft.Rnw
edited introduction and methodology section
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1677 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 8a97252ee7386f3b2504e2505c66c08b6554b3ed

\Abstract{
In the field of probabilistic forecasting classical linear models such as ensemble model output statistics (EMOS, \cite{gneiting2005calibrated}) provide common and widely used tools. This is due to their simplicity and their interpretability. However, the restriction to modeling only linear effects requires further extensions which allow for more flexibility. Generalized additive models for location, scale, and shape (GAMLSS, \cite{stasinopoulos2005}) can also capture non-linear effects as each distribution parameter can be modeled separately by its own generalized additive model (GAM). A different approach that would make it possible to also model non-additive effects and interactions is to apply tree structured models. However, up to now, this class of models has not yet been integrated in probabilistic forecasting as no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. Applying these novel approaches on precipitation data in a mountainous region with a large number of numerical weather prediction quantities illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge/experience from the meteorology or based on a computationally more demanding boosting approach.

@@

\SweaveOpts{eval = TRUE}
\section{Introduction}
A typical application in probabilistic forecasting in meteorology is the analysis of precipitation data. The two main interests lie in estimating the probability of rain as well as predicting the expected amount of precipitation. In this research area various methods have been developed and tested in the last decades. Basic linear models such as ensemble model output statistics (EMOS) were introduced more than a decade ago (\cite{gneiting2005calibrated}) and are still common and widely used. Despite their simplicity and interpretability, the limitation to modeling only linear effects has risen the interest in applying more flexible models. This feature is provided by generalized additive models (GAM) which can also model non-linear effects. However, often only the expectation of the underlying distribution is estimated. To obtain probabilistic predictions, generalized additive models for location, scale, and shape (GAMLSS) offer a framework in which each distribution parameter can be modeled separately by its own GAM. In Figure \ref{Devel_parmod} the development of parametric models starting from simple linear models (LM) and ranging to GAMLSS is illustrated.\\
These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth/homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain/mountainous regions various factors can change rapidly and their effects are difficult to be modeled/represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree structured models. Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model in advance is required. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothened in case of smooth effects while abrupt changes are still captured by the tree structure of the model.\\
For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. But, up to now, no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure \ref{Devel_treeforest}.\\

@@

<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
\end{center}
\endminipage
\vspace{0.0cm}
\minipage{0.25\textwidth}
\begin{center}
\hspace{0.4cm}
LM, GLM 
\end{center}
\endminipage
\hspace{1.7cm}
\minipage{0.25\textwidth}
\begin{center}
GAM
\end{center}
\endminipage
\hspace{1.7cm}
\minipage{0.25\textwidth}
\begin{center}
GAMLSS
\end{center}
\endminipage
\caption{\label{Devel_parmod}Development in the field of parametric modeling}
\end{figure}

@@

\newpage
\section{Methodology}
\label{Methodology}
\subsection{Distributional Fit}
Fitting a distributional model to a set of observations without considering any covariates is a well known procedure which can for example be done by applying the maximum likelihood method. This is also the chosen approach in the methodology described in this section.\\
The goal is to fit a distributional model $D(Y, \theta)$ to a response variable $Y \in \mathcal{Y}$ for a \textit{k}-dimensional vector of distribution parameters $\theta \in \Theta$ in case of a \textit{k}-parametric distribution. The distribution family has to be specified in advance such that a log-likelihood function $\ell(\theta; Y)$ is provided. In our application of precipitation forecasting the response variable, which is the power transformed total precipitation amount within 24 hours, is expected to be normally distribution with mean/location parameter $\mu$ and standard deviation/scale parameter $\sigma$ and left censored with censoring point 0. Therefore, the corresponding log-likelihood function $\ell_{tp}$ with parameter vector $\theta = (\mu, \sigma)$ is
$$
\ell_{tp}(\mu, \sigma; Y) = 
\begin{cases}
    \log\left(\frac{1}{\sigma} \cdot \phi(\frac{Y - \mu}{\sigma}) \right), & \text{if } Y > 0\\
    \vspace{0.1cm}
    \log\left(\Phi(\frac{-\mu}{\sigma})\right), & \text{if } Y = 0
\end{cases}
$$
where $\phi$ and $\Phi$ are the probability density function and the distribution function of the standard normal distribution $\mathcal{N}(0,1)$.\\
With this specification the task of fitting a distributional model turns into the task of determining the distribution parameter $\theta$. This is done by the following maximization.
$$
\hat{\theta} = \max_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)
$$
where $\{y_i\}_{i=1,\ldots,n}$ are the observations of the response variable $Y$.\\
In that way, a whole distribution is specified with all its features, including location, scale and shape. However, fitting one global model to the whole data set might be too much of a generalization and therefore not represent the data and its features in a reasonable way. Moreover, if covariates are available it is desirable to include them into the model as they might provide important information. In our application on precipitation data a wide range of covariates are available from numerical ensemble predictions considering precipitation, temperature, sea level pressure and other meteorological quantities which might be of interest/have an influence on the predictions. In particular, these covariates can be used to separate the data into more homogeneous subgroups. Doing so after fitting a global model to the whole data set and then fitting a local model to each subgroup should improve the model remarkably. This procedure of splitting a data set into subgroups can be done by applying a tree algorithm.

\subsection{Distributional Tree}
After having fit a global distributional model $D(Y, \bold{\hat{\theta}})$ the idea is now to use the information provided by this model together with a set of covariates $\bold{Z} = (Z_1, \ldots, Z_m) \in \mathcal{Z}$ to decide if the data set should be split into subgroups. In particular, the information provided by the fitted model are the scores 
$$
s(\hat{\theta}, y_i) = \frac{\partial \ell}{\partial \theta}(\hat{\theta}; y_i).
$$ 
These values express how well the model fits the data regarding each parameter. Ideally they should fluctuate randomly around zero, similar to residuals in ordinary least squares estimations, but in this case there is a score value for each estimated parameter $\hat{\theta}_j \in (\hat{\theta}_1, \ldots, \hat{\theta}_k) = \bold{\hat{\theta}}$ and each observation.\\
Statistical tests are applied to find out whether any significant connection/dependency between the scores and each of the partitioning variables $Z_l \in \bold{Z} = (Z_1, \ldots, Z_m)$ is given. If so, the data set is split into subgroups based on the values of the partitioning variable that shows the highest connection/dependency.\\
In a next step a local distributional model is fit in each of the subgroups. Repeating this procedure of fitting a model, applying statistical tests to evaluate the goodness of fit and splitting the data set depending on the test results leads to a tree structure. In that way the learning data set is separated into \textit{B} disjoint segments such that
$$
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b
$$
where each segment is represented by a terminal node of the tree.\\

@@

\subsection{Distributional Forest}
A distributional forest consists of an ensemble of distributional trees $\{t_s\}_{s=1,\ldots,T}$. Each of these trees is built on a different data set which can either be a bootstrap or a subset of the original data set. Moreover, in each node a subset of the partitioning variables is chosen randomly as candidates for splitting variables. In that way the correlation among the resulting trees is reduced and therefore also the variance of the model.\\
\\
Forest models differ mainly in the way how the trees are combined to make predictions. In this case the forest itself is only used to evaluate neighborhood relationships among the observations which can then be included in the fitting process.\\
Contrary to distributional trees, in distributional forests the whole learning data set can be used to fit the distributional model and specify the predicted distribution for a new observation. This difference is represented in the weight vector, as the tree weights $\bold{\omega}^{tree}$ containing only zeros and ones are now replaced by integer valued forest weights \\
$\bold{\omega}^{forest} = (\omega^{forest}_1, \ldots, \omega^{forest}_n)$.\\
For a new observation $\bold{z}^{\ast} = (z^{\ast}_1, \ldots, z^{\ast}_m) $ this set of weights is obtained by the following procedure: For each observation $(y_j,\bold{z}_j)$ in the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ a weight is calculated by counting the number of trees in which it ends up in the same terminal node as the new observation $\bold{z}^{\ast}$. 
$$
\omega^{forest}_i(\bold{z}^{\ast}) = \sum_{s=1}^T \sum_{b=1}^{B^s} I((\bold{z}_i \in \mathcal{B}^s_b) \land (\bold{z}^{\ast} \in \mathcal{B}^s_b))
$$
where $T$ is the number of trees of the forest, $B^s$ is the number of terminal nodes of the \textit{s}-th tree and $\mathcal{B}^s_b$ represents the \textit{b}-th terminal node of the \textit{s}-th tree.\\
This individual set of weights $\bold{\omega}^{forest}(\bold{z}^{\ast}) = (\omega^{forest}_1(\bold{z}^{\ast}), \ldots, \omega^{forest}_n(\bold{z}^{\ast}))$ for the new observation $\bold{z}^{\ast}$ can now be included in the estimation process of the distribution parameters and in that way, leads to its individual parameter vector $\hat{\theta}(\bold{z}^{\ast})$.
$$
\hat{\theta}(\bold{z}^{\ast}) =  \max_{\theta \in \Theta} \sum_{i=1}^n \omega^{forest}_i(\bold{z}^{\ast}) \cdot \ell(\theta; y_i)
$$
While in a distributional tree the same distribution is specified for all observations being in the same node a distributional forest predicts and fits a different distribution for each different observation which is due to the individual weights calculated separately for each observation.

\newpage
\section{Simulation Study}
Within this simulation study the performance of distributional trees and forests is tested and compared to other models, showing situations where one or the other method is more advantageous than others. By doing so the following hypotheses are investigated:
\begin{itemize}%[leftmargin=0.5cm]
\item When applying a tree algorithm the use of a fitting function which returns scores for each distribution parameter is advantageous to find correct splits. In particular it is crucial to find splits in parameters other than the location parameter. Hence a distributional forest consisting of distributional trees leads to better results than a simple forest considering only the mean.
\item In the case of smooth parameter functions without any abrupt changes / non-additive effects a smooth parametric model such as GAMLSS is clearly the first choice while a stepfunction can be modeled best by applying a tree-structured model. Even though for both of these extreme situations a specifically advantageous method is provided, a forest model can be a good compromise and performs well in a wide range of settings including the two mentioned before. (Therefore, a forest is a recommendable choice particularly when no information about the type of parameter function is provided.)
\item While specifying interactions and selecting covariates is necessary to achieve reasonable results in parametric models such as GAMLSS, this is done automatically within trees and forests. For that reason, distributional trees and forests provide a framework where no additional effort is required for the model specification. This feature plays to its strength especially in case of a high number of covariates and no information about their effects on the response.
\end{itemize}

@@

\begin{figure}[t!]
\begin{subfigure}{0.6\textwidth}
<<sim_rmse_exp, fig=TRUE, echo=FALSE>>=
plot_rmse(simres, type = "exp")
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.6\textwidth}
<<sim_rmse_par, fig=TRUE, echo=FALSE>>=
plot_rmse(simres, type = "par")
@
\end{subfigure}
\caption{\label{plot_sim_rmse}Results of the simulation study (RMSE)\\
left: continuous lines represent the difference between the true and the estimated expected value, dotted lines represent the difference between the observations and the estimated expected value\\
right: continuous lines represent the location parameter, dotted lines the scale parameter}
\end{figure}

@@

\SweaveOpts{eval = TRUE}
\newpage
\newpage
\section{Probabilistic Forecasting of Precipitation}
As explained in the introduction distributional trees and forests provide a new approach in probabilistic forecasting which combines advantages of parametric models that are already used in this field and tree structured models which provide features that could be particularly valuable in probabilistic forecasting but have not yet been applied in this kind of settings. To show their strengths in practice, these new methods are now applied to probabilistic precipitation forecasts in a mountainous region based on a large number of numerical weather prediction quantities. In particular, the data set consists of daily 24h precipitation sums for the month of July from 28 years, starting in 1985 until 2012, both included. The observations were measured at 95 (?) stations all over Tyrol (and regions close to the frontier). The considered data is provided by ..... 
Next to the variable of interest, the measured total precipitation within 24 hours of each day (measured from 6 UTC to 6 UTC of the following day), numerical forecasts from the ECMWF are used as covariates. These forecasts are ensemble predictions consisting of ... individual forecasts based on perturbed initial conditions.\\
In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since each forecast variable consists of an ensemble of individual predictions the possible variations include the mean, the standard deviation, the minimum and the maximum over all these individual predictions, noted as ensemble mean, ensemble std dev, ensemble minimum and ensemble maximum. All of these values can be calculated on the whole day (6--30 UTC) or a 6h time window of the day.\\
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std dev of mean/minimal/maximal values respectively.\\
Table \ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82 covariates are available in this setting.\\
The response variable, total precipitation (tp), is power transformed by a factor of 1.6. This is done due to experience and knowledge about possible distributions of precipitation measurements. The resulting transformed variable is expected to be normally distributed and left censored at censoring point 0.

@@

\begin{figure}[t!]
\begin{center}
<<rain_Axams_pred, fig=TRUE, echo=FALSE>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday, dayending), ylab = "Density", 
     xlab = "Total precipitation (power transformed)",
     ylim = c(0,max(y1, y2, y3, pm1, pm2, pm3) + 0.01),
     xlim = c(-1.5,8))

@@

\begin{figure}[t!]
\begin{center}
\begin{subfigure}{0.8\textwidth}
<<rain_axams_pit_oosample, fig=TRUE, echo=FALSE>>=

@@

@
\caption{\label{map}Map of all observations with corresponding CRPS skills score for predictions}
\end{center}
\end{figure}

>>>>

\Abstract{
In the field of probabilistic forecasting linear models such as ensemble model output statistics (EMOS, \cite{gneiting2005calibrated}) provide common and widely used tools. This is due to their simplicity and their interpretability. However, the restriction to modeling only linear effects requires further extensions which allow for more flexibility. Generalized additive models for location, scale, and shape (GAMLSS, \cite{stasinopoulos2005}) can also capture non-linear effects as each distribution parameter can be modeled separately by its own generalized additive model (GAM). A different approach that would make it possible to also model non-additive effects and interactions is to apply tree structured models such as regression trees (\cite{loh2011classification}) or random forests (\cite{breiman2001random}). However, up to now, this class of models has not yet been integrated in probabilistic forecasting as no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. Applying these novel approaches on precipitation data in a mountainous region with a large number of numerical weather prediction quantities illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge/experience from the meteorology or based on a computationally more demanding boosting approach.

@@

\SweaveOpts{eval = TRUE}
\section{Introduction}
A typical application in probabilistic forecasting in meteorology is the analysis of precipitation data. The two main interests lie in estimating the probability of rain as well as predicting the expected amount of precipitation. In this research area various methods have been developed and tested in the last decades. (Multiple) Linear models such as ensemble model output statistics (EMOS) were introduced more than a decade ago (\cite{gneiting2005calibrated}) and are still common and widely used. Despite their simplicity and interpretability, the limitation to modeling only linear effects has risen the interest in applying more flexible models. This feature is provided by generalized additive models (GAM) which can also model non-linear effects. However, often only the expectation of the underlying distribution is estimated. To obtain probabilistic predictions, generalized additive models for location, scale, and shape (GAMLSS, \cite{stasinopoulos2005}) offer a framework in which each distribution parameter can be modeled separately by its own GAM. In Figure \ref{Devel_parmod} the development of parametric models starting from simple linear models (LM) and ranging to GAMLSS is illustrated.\\
These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth/homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain/mountainous regions various factors can change rapidly and their effects are difficult to be modeled/represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree structured models such as regression trees (\cite{loh2011classification}) or random forests (\cite{breiman2001random}). Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model in advance is required. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothed out in case of smooth effects while abrupt changes are still captured by the tree structure of the model.\\
For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. But, up to now, no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure \ref{Devel_treeforest}.\\
After explaining the novel methodology in section \ref{Methodology} of this paper and presenting results from a simulation study in section \ref{Simulation} the focus will then be put on an application in the field of probabilistic forecasting in section \ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration and based on the data/information from 25 years probabilistic predictions are made for the 3 successive years. A brief outlook on the results of this setting is given by Figure \ref{axams_onestation}. For the 24th of July of 2010, 2011 and 2012 the density functions of the estimated distributions for this day of all three prediction years are plotted together with the corresponding observations. The estimated distribution parameters for the 24th of July of 2010 are comparably high with location $3.146$ and scale $1.902$ which leads to a well fitting predicted expectation of $3.185$ close to the observed value of $3.001$. Also for 2011 and 2012 the estimated expected values of 1.085 and 0.483 provide reasonable predictions as the observed values are 1.121 and 0 respectively.\\
Next to the expected values the predictions provided by these distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis.

\begin{figure}[t!]
\begin{center}
<<rain_Axams_pred, fig=TRUE, echo=FALSE, width=6.5, height=4.5>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday, dayending), ylab = "Density", 
     xlab = "Total precipitation (power transformed)",
     ylim = c(0,max(y1, y2, y3, pm1, pm2, pm3) + 0.01),
     xlim = c(-1.5,8))

    # add labels
    text(x = -0.8, y = lh1, labels = "2010", col = "red", cex = 0.8)
    text(x = -0.8, y = lh2, labels = "2011", col = "blue", cex = 0.8)
    text(x = -0.8, y = lh3, labels = "2012", col = "darkgreen", cex = 0.8)
@
\caption{\label{axams_onestation}Predictions at station Axams for one day in July of 2010, 2011 and 2012.}
\end{center}
\end{figure}

@@

<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
\end{center}
\endminipage
\vspace{0.0cm}
\minipage{0.25\textwidth}
\begin{center}
\hspace{0.4cm}
LM, GLM 
\end{center}
\endminipage
\hspace{1.7cm}
\minipage{0.25\textwidth}
\begin{center}
GAM
\end{center}
\endminipage
\hspace{1.7cm}
\minipage{0.25\textwidth}
\begin{center}
GAMLSS
\end{center}
\endminipage
\caption{\label{Devel_parmod}Development in the field of parametric modeling.}
\end{figure}

@@

\newpage
\section{Methodology}
\label{Methodology}

\subsection{Motivation (Review/Problems)}
As illustrated in Figure \ref{Devel_parmod} a lot of developement has been made in the field of parametric modeling and a wide range of tools can be applied in research areas such as probabilistic forecasting. But as mentioned before, further features such as the ability to capture non-additive effects or interactions are desirable in many applications which demands for more flexible models. Since tree structured models provide these required features a combination of them with parametric models can offer a framework that is able to deal with a wide range of different modeling situations. In this section we are going to introduce the novel methodology which embeds parametric modeling into the idea of tree structered models step by step starting with a simple distributional fit in subsection \ref{distfit}, then building a distributional tree in subsection \ref{disttree} and finally a distributional forest in subsection \ref{distforest}.

\subsection{Distributional Fit}
\label{distfit}
Fitting a distributional model to a set of observations without considering any covariates is a well known procedure which can for example be done by applying the maximum likelihood method. This is also the chosen approach in the methodology described in this section.\\
The goal is to fit a distributional model $D(Y, \theta)$ to a response variable $Y \in \mathcal{Y}$ for a \textit{k}-dimensional vector of distribution parameters $\theta \in \Theta$ in case of a \textit{k}-parametric distribution. The distribution family has to be specified in advance such that a log-likelihood function $\ell(\theta; Y)$ is provided. Using the GAMLSS framework (\cite{stasinopoulos2005}) a wide range of distributions can be selected where each distribution parameter can be specified separatly. This enables the model to deal with various features that can usually be very challenging in model fitting such as censoring, truncation or heavy tails.\\
In our application of precipitation forecasting the response variable, which is the power transformed total precipitation amount within 24 hours, is expected to be normally distributed with mean/location parameter $\mu$ and standard deviation/scale parameter $\sigma$ and left censored with censoring point 0. Therefore, the corresponding log-likelihood function with parameter vector $\theta = (\mu, \sigma)$ is
$$
\ell(\mu, \sigma; Y) = 
\begin{cases}
    \log\left(\frac{1}{\sigma} \cdot \phi(\frac{Y - \mu}{\sigma}) \right), & \text{if } Y > 0\\
    \vspace{0.1cm}
    \log\left(\Phi(\frac{-\mu}{\sigma})\right), & \text{if } Y = 0
\end{cases}
$$
where $\phi$ and $\Phi$ are the probability density function and the distribution function of the standard normal distribution $\mathcal{N}(0,1)$. Next to the censored normal distribution the censored logistic distribution is also often chosen for the modeling of precipitation amounts. In \cite{Gebetsberger2017} this choice has been made in order to be able to deal with heavy tails which often appear in precipitation data. Another frequent application of probabilistic forecasting in meteorology is the modeling of temperature. In this case a normal distribution is often selected.\\
With the specification of the distribution family and the log-likelihood function the task of fitting a distributional model turns into the task of determining the distribution parameter $\theta$. This is done by the following maximization.
$$
\hat{\theta} = \max_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)
$$
where $\{y_i\}_{i=1,\ldots,n}$ are the observations of the response variable $Y$.\\
In that way, a whole distribution is specified with all its features, including location, scale and shape. However, fitting one global model to the whole data set might be too much of a generalization and therefore not represent the data and its features in a reasonable way. Moreover, if covariates are available it is desirable to include them into the model as they might provide important information. In our application on precipitation data a wide range of covariates are available from numerical ensemble predictions considering precipitation, temperature, sea level pressure and other meteorological quantities which might be of interest/have an influence on the predictions. \\
A common way of including these covariates is to consider them as regressors in a distributional model such that each parameter $\theta$ depends on covariates $\bold{X} = (X_1,\ldots,X_m)$. This could for example be done by a linear function $g$. In that way the fitted model would be specified as $D(Y, \theta = g(\bold{X}))$. In order to avoid the restriction to linear effects one might also use a set of smooth functions $\{f_i\}_{i = 1,\ldots,p}$ to specify a more flexible model $D(Y, \theta = f_1(\bold{X}) + \ldots + f_p(\bold{X}))$ applying a GAM structure. Another approach to distributional regression is proposed by \cite{Klein2015} where each parameter of a possibly complex distribution can be modeled by an additive term consisting of a variety of different functional effect types. These include non-linear effects, spatial effects, random coefficients and interaction surfaces which extends the flexibility compared to parametric models such as GAMLSS modeling only smooth effects.\\
However, the restriction to additive models remains and therefore a different approach of how to include covariates that also allows for non-additive effects and interactions is now considered. In particular, the covariates can be used to separate the data into more homogeneous subgroups. Doing so after fitting a global model to the whole data set (without considering any covariates) and then fitting a local model to each subgroup should improve the model remarkably. This procedure of splitting a data set into subgroups can be done by applying a tree algorithm.\\

\subsection{Distributional Tree}
\label{disttree}
After having fit a global distributional model $D(Y, \bold{\hat{\theta}})$ the idea is now to use the information provided by this model together with a set of covariates $\bold{Z} = (Z_1, \ldots, Z_m) \in \mathcal{Z}$ (from now on denoted by $\bold{Z}$ instead of $\bold{X}$ to indicate that they are considered as split variables only and not as regressors) to decide if the data set should be split into subgroups. In particular, the information provided by the fitted model are the scores 
$$
s(\hat{\theta}, y_i) = \frac{\partial \ell}{\partial \theta}(\hat{\theta}; y_i).
$$ 
These values express how well the model fits the data regarding each parameter. Ideally they should fluctuate randomly around zero, similar to residuals in ordinary least squares estimations, but in this case there is a score value for each estimated parameter $\hat{\theta}_j \in (\hat{\theta}_1, \ldots, \hat{\theta}_k) = \bold{\hat{\theta}}$ and each observation.\\
Statistical tests are applied to find out whether any significant connection/dependency between the scores and each of the partitioning variables $Z_l \in \bold{Z} = (Z_1, \ldots, Z_m)$ is given. If so, the data set is split into subgroups based on the values of the partitioning variable that shows the highest connection/dependency.\\
In a next step a local distributional model is fit in each of the subgroups. Repeating this procedure of fitting a model, applying statistical tests to evaluate the goodness of fit and splitting the data set depending on the test results leads to a tree structure. In that way the learning data set is separated into \textit{B} disjoint segments such that
$$
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b
$$
where each segment is represented by a terminal node of the tree.\\

@@

\subsection{Distributional Forest}
\label{distforest}
A distributional forest consists of an ensemble of distributional trees $\{t_s\}_{s=1,\ldots,T}$. Each of these trees is built on a different data set which can either be a bootstrap or a subset of the original data set. Moreover, in each node a subset of the partitioning variables is chosen randomly as candidates for splitting variables. In that way the correlation among the resulting trees is reduced and therefore also the variance of the model (\cite{breiman2001random}).\\
\\
Forest models differ mainly in the way how the trees are combined to make predictions. In this case the forest itself is only used to evaluate neighborhood relationships among the observations which can then be included in the fitting process.\\
Contrary to distributional trees, in distributional forests the whole learning data set can be used to fit the distributional model and specify the predicted distribution for a new observation. This difference is represented in the weight vector, as the tree weights $\bold{\omega}^{tree}$ containing only zeros and ones are now replaced by integer valued forest weights \\
$\bold{\omega}^{forest} = (\omega^{forest}_1, \ldots, \omega^{forest}_n)$.\\
For a new observation $\bold{z}^{\ast} = (z^{\ast}_1, \ldots, z^{\ast}_m) $ this set of weights is obtained by the following procedure: For each observation $(y_j,\bold{z}_j)$ in the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ a weight is calculated by counting the number of trees in which it ends up in the same terminal node as the new observation $\bold{z}^{\ast}$. 
$$
\omega^{forest}_i(\bold{z}^{\ast}) = \sum_{s=1}^T \sum_{b=1}^{B^s} I((\bold{z}_i \in \mathcal{B}^s_b) \land (\bold{z}^{\ast} \in \mathcal{B}^s_b))
$$
where $T$ is the number of trees of the forest, $B^s$ is the number of terminal nodes of the \textit{s}-th tree and $\mathcal{B}^s_b$ represents the \textit{b}-th terminal node of the \textit{s}-th tree.\\
This individual set of weights $\bold{\omega}^{forest}(\bold{z}^{\ast}) = (\omega^{forest}_1(\bold{z}^{\ast}), \ldots, \omega^{forest}_n(\bold{z}^{\ast}))$ for the new observation $\bold{z}^{\ast}$ can now be included in the estimation process of the distribution parameters and in that way, leads to its individual parameter vector $\hat{\theta}(\bold{z}^{\ast})$.
$$
\hat{\theta}(\bold{z}^{\ast}) =  \max_{\theta \in \Theta} \sum_{i=1}^n \omega^{forest}_i(\bold{z}^{\ast}) \cdot \ell(\theta; y_i)
$$
While in a distributional tree the same distribution is specified for all observations being in the same node a distributional forest predicts and fits a different distribution for each different observation which is due to the individual weights calculated separately for each observation.

\newpage
\section{Simulation Study}
\label{Simulation}
Within this simulation study the performance of distributional trees and forests is tested and compared to other models, showing situations where one or the other method is more advantageous than others. By doing so the following hypotheses are investigated:
\begin{itemize}%[leftmargin=0.5cm]
\item When applying a tree algorithm the use of a fitting function which returns scores for each distribution parameter is advantageous to find correct splits. In particular it is crucial to find splits in parameters other than the location parameter. Hence a distributional forest consisting of distributional trees leads to better results than a simple forest considering only the mean.
\item In the case of smooth parameter functions without any abrupt changes / non-additive effects a smooth parametric model such as GAMLSS is clearly the first choice while a stepfunction can be modeled best by applying a tree-structured model. Even though for both of these extreme situations a specifically advantageous method is provided, a forest model can be a good compromise and performs well in a wide range of settings including the two mentioned before. (Therefore, a forest is a recommendable choice particularly when no information about the type of parameter function is provided.)
\item While specifying interactions and selecting covariates is necessary to achieve reasonable results in parametric models such as GAMLSS, this is done automatically within trees and forests. For that reason, distributional trees and forests provide a framework where no additional effort is required for the model specification. This feature plays to its strength especially in case of a high number of covariates and no information about their effects on the response.
\end{itemize}

@@

\begin{figure}[t!]
\begin{subfigure}{0.6\textwidth}
<<sim_rmse_exp, fig=TRUE, echo=FALSE>>=
plot_rmse(simres, type = "exp")
@
\end{subfigure}
\hspace{-1cm}
\begin{subfigure}{0.6\textwidth}
<<sim_rmse_par, fig=TRUE, echo=FALSE>>=
plot_rmse(simres, type = "par")
@
\end{subfigure}
\caption{\label{plot_sim_rmse}Results of the simulation study (RMSE).\\
left: continuous lines represent the difference between the true and the estimated expected value, dotted lines represent the difference between the observations and the estimated expected value\\
right: continuous lines represent the location parameter, dotted lines the scale parameter}
\end{figure}

@@

\SweaveOpts{eval = TRUE}
\newpage
\newpage
\section{Probabilistic Forecasting of Precipitation}
\label{ProbabilisticForecasting}
As explained in the introduction distributional trees and forests provide a new approach in probabilistic forecasting which combines advantages of parametric models that are already used in this field and tree structured models which provide features that could be particularly valuable in probabilistic forecasting but have not yet been applied in this kind of settings. To show their strengths in practice, these new methods are now applied to probabilistic precipitation forecasts in a mountainous region based on a large number of numerical weather prediction quantities. In particular, the data set consists of daily 24h precipitation sums for the month of July from 28 years, starting in 1985 until 2012, both included. The observations were measured at 95 (?) stations all over Tyrol (and regions close to the frontier). The considered data is provided by ..... 
Next to the variable of interest, the measured total precipitation within 24 hours of each day (measured from 6 UTC to 6 UTC of the following day), numerical forecasts from the ECMWF are used as covariates. These forecasts are ensemble predictions consisting of ... individual forecasts based on perturbed initial conditions.\\
In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since each forecast variable consists of an ensemble of individual predictions the possible variations include the mean, the standard deviation, the minimum and the maximum over all these individual predictions, noted as ensemble mean, ensemble std dev, ensemble minimum and ensemble maximum. All of these values can be calculated on the whole day (6--30 UTC) or a 6h time window of the day.\\
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std dev of mean/minimal/maximal values respectively.\\
Table \ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82 covariates are available in this setting.\\
The response variable, total precipitation (tp), is power transformed by a factor of 1.6. This is done due to experience and knowledge about possible distributions of precipitation measurements. The resulting transformed variable is expected to be normally distributed and left censored at censoring point 0.

@@

\begin{figure}[t!]
\begin{center}
<<rain_Axams_pred, fig=TRUE, echo=FALSE, width=6.5, height=4.5>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday, dayending), ylab = "Density", 
     xlab = "Total precipitation (power transformed)",
     ylim = c(0,max(y1, y2, y3, pm1, pm2, pm3) + 0.01),
     xlim = c(-1.5,8))

@@

\begin{figure}[t!]
\begin{center}
\begin{subfigure}{0.8\textwidth}
<<rain_axams_pit_oosample, fig=TRUE, echo=FALSE, width = 6.5>>=

@@

@
\caption{\label{map}Map of all observations with corresponding CRPS skills score for predictions.}
\end{center}
\end{figure}
